---
title: "Supervised machine learning"
author: Pablo Barbera
date: August 9, 2018
output: html_document
---

### Regularized regression

To learn how to do supervised machine learning applied to social media text, we will use a random sample of nearly 5,000 tweets mentioning the names of the candidates to the 2014 EP elections in the UK. We will be analyzing variable named `communication`, which indicates whether each tweet was hand-coded as being __engaging__ (a tweet that tries to engage with the audience of the account) or __broadcasting__ (just sending a message, without trying to elicit a response).

The source of the dataset is an article co-authored with Yannis Theocharis, Zoltan Fazekas, and Sebastian Popa, published in the Journal of Communication. The link is [here](http://onlinelibrary.wiley.com/doi/10.1111/jcom.12259/abstract). Our goal was to understand to what extent candidates are not engaging voters on Twitter because they're exposed to mostly impolite messages.

Let's start by reading the dataset and creating a dummy variable indicating whether each tweet is engaging

```{r}
library(quanteda)
tweets <- read.csv("~/data/UK-tweets.csv", stringsAsFactors=F)
tweets$engaging <- ifelse(tweets$communication=="engaging", 1, 0)
tweets <- tweets[!is.na(tweets$engaging),]
```

We'll do some cleaning as well -- substituting handles with @. Why? We want to provent overfitting.

```{r}
tweets$text <- gsub('@[0-9_A-Za-z]+', '@', tweets$text)
```

As we discussed earlier today, before we can do any type of automated text analysis, we will need to go through several "preprocessing" steps before it can be passed to a statistical model. We'll use the quanteda package quanteda here.

```{r}
twcorpus <- corpus(tweets$text)
summary(twcorpus)
```

We can then convert a corpus into a document-feature matrix using the dfm function. We will then trim it in order to keep only tokens that appear in 2 or more tweets. Note that we keep punctuation -- it turns out it can be quite informative.


```{r}
twdfm <- dfm(twcorpus, remove=stopwords("english"), remove_url=TRUE, 
             ngrams=1:2, verbose=TRUE)
twdfm <- dfm_trim(twdfm, min_docfreq = 2, verbose=TRUE)
```

Note that other preprocessing options are:

- remove_numbers
- remove_punct
- remove_twitter
- remove_symbols
- remove_separators

You can read more in the `dfm` and `tokens` help pages

Once we have the DFM, we split it into training and test set. We'll go with 80% training and 20% set. Note the use of a random seed to make sure our results are replicable.

```{r}
set.seed(123)
training <- sample(1:nrow(tweets), floor(.80 * nrow(tweets)))
test <- (1:nrow(tweets))[1:nrow(tweets) %in% training == FALSE]
```

Our first step is to train the classifier using cross-validation. There are many packages in R to run machine learning models. For regularized regression, glmnet is in my opinion the best. It's much faster than caret or mlr (in my experience at least), and it has cross-validation already built-in, so we don't need to code it from scratch. We'll start with a ridge regression:

```{r}
library(glmnet)
require(doMC)
registerDoMC(cores=3)
ridge <- cv.glmnet(twdfm[training,], tweets$engaging[training], 
	family="binomial", alpha=0, nfolds=5, parallel=TRUE, intercept=TRUE,
	type.measure="class")
plot(ridge)
```

We can now compute the performance metrics on the test set.
```{r}
## function to compute accuracy
accuracy <- function(ypred, y){
	tab <- table(ypred, y)
	return(sum(diag(tab))/sum(tab))
}
# function to compute precision
precision <- function(ypred, y){
	tab <- table(ypred, y)
	return((tab[2,2])/(tab[2,1]+tab[2,2]))
}
# function to compute recall
recall <- function(ypred, y){
	tab <- table(ypred, y)
	return(tab[2,2]/(tab[1,2]+tab[2,2]))
}
# computing predicted values
preds <- predict(ridge, twdfm[test,], type="class")
# confusion matrix
table(preds, tweets$engaging[test])
# performance metrics
accuracy(preds, tweets$engaging[test])
precision(preds==1, tweets$engaging[test]==1)
recall(preds==1, tweets$engaging[test]==1)
precision(preds==0, tweets$engaging[test]==0)
recall(preds==0, tweets$engaging[test]==0)
```

Something that is often very useful is to look at the actual estimated coefficients and see which of these have the highest or lowest values:

```{r}
# from the different values of lambda, let's pick the highest one that is
# within one standard error of the best one (why? see "one-standard-error"
# rule -- maximizes parsimony)
best.lambda <- which(ridge$lambda==ridge$lambda.1se)
beta <- ridge$glmnet.fit$beta[,best.lambda]
head(beta)

## identifying predictive features
df <- data.frame(coef = as.numeric(beta),
				word = names(beta), stringsAsFactors=F)

df <- df[order(df$coef),]
head(df[,c("coef", "word")], n=30)
paste(df$word[1:30], collapse=", ")
df <- df[order(df$coef, decreasing=TRUE),]
head(df[,c("coef", "word")], n=30)
paste(df$word[1:30], collapse=", ")
```

We can easily modify our code to experiment with Lasso or ElasticNet models:

```{r}
lasso <- cv.glmnet(twdfm[training,], tweets$engaging[training], 
	family="binomial", alpha=1, nfolds=5, parallel=TRUE, intercept=TRUE,
	type.measure="class")

```

```{r}
# computing predicted values
preds <- predict(lasso, twdfm[test,], type="class")
# confusion matrix
table(preds, tweets$engaging[test])
# performance metrics (slightly better!)
accuracy(preds, tweets$engaging[test])
precision(preds==1, tweets$engaging[test]==1)
recall(preds==1, tweets$engaging[test]==1)
precision(preds==0, tweets$engaging[test]==0)
recall(preds==0, tweets$engaging[test]==0)

```

```{r}
best.lambda <- which(lasso$lambda==lasso$lambda.1se)
beta <- lasso$glmnet.fit$beta[,best.lambda]
head(beta)

## identifying predictive features
df <- data.frame(coef = as.numeric(beta),
				word = names(beta), stringsAsFactors=F)

df <- df[order(df$coef),]
head(df[,c("coef", "word")], n=30)

df <- df[order(df$coef, decreasing=TRUE),]
head(df[,c("coef", "word")], n=30)
```

We now see that the coefficients for some features actually became zero.

```{r}
enet <- cv.glmnet(twdfm[training,], tweets$engaging[training], 
	family="binomial", alpha=0.50, nfolds=5, parallel=TRUE, intercept=TRUE,
	type.measure="class")
# NOTE: this will not cross-validate across values of alpha

# computing predicted values
preds <- predict(enet, twdfm[test,], type="class")
# confusion matrix
table(preds, tweets$engaging[test])
# performance metrics
accuracy(preds, tweets$engaging[test])
precision(preds==1, tweets$engaging[test]==1)
recall(preds==1, tweets$engaging[test]==1)
precision(preds==0, tweets$engaging[test]==0)
recall(preds==0, tweets$engaging[test]==0)


best.lambda <- which(enet$lambda==enet$lambda.1se)
beta <- enet$glmnet.fit$beta[,best.lambda]
head(beta)

## identifying predictive features
df <- data.frame(coef = as.numeric(beta),
				word = names(beta), stringsAsFactors=F)

df <- df[order(df$coef),]
head(df[,c("coef", "word")], n=30)

df <- df[order(df$coef, decreasing=TRUE),]
head(df[,c("coef", "word")], n=30)

```

