---
title: "Querying an SQL database"
author: Pablo Barbera
date: August 11, 2017
output: html_document
---

## Creating an SQL database

For this first part of the class, we'll be working with a SQLite database, which is self-contained in a file within our hard drive, without any need to set up a server. The dataset we will work with is all Facebook posts by Members of the U.S. Congress in 2017.

We'll start by connecting to the database and adding the first table, with information at the member of congress level. 

```{r}
library(DBI)
```

```{r, eval=FALSE}
# create database: this will create a file in our hard drive
db <- dbConnect(RSQLite::SQLite(), "facebook-db.sqlite")

# reading the first file
congress <- read.csv("../data/congress-facebook-2017.csv",
	stringsAsFactors=F)

# adding first table: user-level data
dbWriteTable(db, "congress", congress)

# testing that it works with a simple query
dbListFields(db, "congress")
dbGetQuery(db, 'SELECT * FROM congress LIMIT 5')
```

Note that we have used `dbWriteTable` to add the data frame to our database, and then check that it all worked with `dbListFields` and `dbGetQuery`. As we will see in a second, this second function is what we will use to query the database, and it will always return a table.

We now switch to the second table. However, the files are too big to open them all in memory. Instead, we will open them one by one, and then __append__ them to the table. Let's see how that works...

```{r, eval=FALSE}

# now adding FB posts
# file is in .zip file, so we unzip it
unzip("../data/facebook-2017.zip")
fls <- list.files("../facebook-2017", full.names=TRUE)

for (f in fls){
	message(f)
	# reading file
	fb <- read.csv(f, stringsAsFactors=F)
	# adding name of file as screen_name (we will use to link tables)
	fb$screen_name <- gsub("../facebook-2017/|.csv", "", f)
	# converting date to R format (so that we can query it later)
	fb$date <- as.character(netdemR::formatFbDate(fb$created_time, "date"))
	fb$datetime <- as.character(netdemR::formatFbDate(fb$created_time, "datetime"))
	# adding to table in SQL database
	dbWriteTable(db, "posts", fb, append=TRUE)
}

# zip it back to save space in our hard drive
zip("../data/facebook-2017.zip", fls)

# testing that it works
dbListFields(db, "posts")
dbGetQuery(db, 'SELECT * FROM posts LIMIT 5')

# what if we make a mistake and want to remove the table?
# dbRemoveTable(db, "posts")

# and we close the connection for now
dbDisconnect(db)

```

## Querying an SQL database

Now that we have our tables in the database, let's see how we can query them. First we connect using `dbConnect` and then we type our query with `dbGetQuery`. Generally, we will write the query __clauses__ in capital letters and the column names in lowercase.

```{r}
db <- dbConnect(RSQLite::SQLite(), "facebook-db.sqlite")
test <- dbGetQuery(db, 'SELECT * FROM congress LIMIT 5')
str(test)
```

RMarkdown actually allows us to write the SQL query directly as a chunk of code:

```{sql, connection=db}
SELECT * FROM congress LIMIT 5
```

This is the list of query clauses that we will use to work with the database:

- __SELECT__: determines which columns to include in the query's result set

- __FROM__: identifies the tables from which to draw data and how tables should be joined

- __LIMIT__: in combination with SELECT, limits the number of records returned

- __WHERE__: filter out unwanted data

- __GROUP BY__: groups rows together by common column values

- __ORDER BY__: sorts the rows in the final result set by one or more columns


Let's start with some examples of __SELECT__:

```{r}
# querying just one column
dbGetQuery(db, "SELECT name FROM congress LIMIT 10")

# multiple columns
dbGetQuery(db, "SELECT name, party FROM congress LIMIT 10")

# adding expressions
dbGetQuery(db, "SELECT from_name, likes_count/comments_count, UPPER(type) FROM posts LIMIT 10")

# adding aliases to the new columns
dbGetQuery(db, "SELECT from_name, likes_count/comments_count AS like_ratio FROM posts LIMIT 10")

# often good idea to write query in multiple lines so that it's easier to read it:
dbGetQuery(db, "SELECT from_name, 
           likes_count/comments_count AS like_ratio 
           FROM posts 
           LIMIT 10")

# selecting unique values from a column
dbGetQuery(db, "SELECT DISTINCT from_name 
           FROM posts 
           LIMIT 10")

```

Now some examples of how we can use __WHERE__ to subset the database based on values of specific columns:

```{r}
# selecting based on values of a column
dbGetQuery(db, "SELECT name, party 
           FROM congress
           WHERE party = 'Republican'
           LIMIT 10")

# greater than, working with dates
dbGetQuery(db, "SELECT from_name, type, date 
           FROM posts
           WHERE date > '2017-01-01'
           LIMIT 10")

# AND operator
dbGetQuery(db, "SELECT from_name, type, date, likes_count 
           FROM posts
           WHERE date > '2017-06-01' AND type != 'photo' AND likes_count > 500
           LIMIT 10")

# OR operator
dbGetQuery(db, "SELECT from_name, type, date, comments_count 
           FROM posts
           WHERE comments_count>100 AND (type = 'photo' OR type = 'video')
           LIMIT 10")

# membership
dbGetQuery(db, "SELECT from_name, type, date, comments_count 
           FROM posts
           WHERE type IN ('video', 'event')
           LIMIT 10")

# matching conditions:
# _ matches exactly one character
# % matches any number of characters
dbGetQuery(db, "SELECT from_name, type, date, comments_count 
           FROM posts
           WHERE date LIKE '2017-01-__'
           LIMIT 10")

dbGetQuery(db, "SELECT from_name, type, date, comments_count 
           FROM posts
           WHERE date LIKE '2017-%'
           LIMIT 10")

# SQLite does not have regular expressions, but we can get creative...
dbGetQuery(db, "SELECT message
           FROM posts
           WHERE message LIKE '%hungary%'
           LIMIT 1")

```


Very useful for us -- __GROUP BY__ allows us to aggregate results based on values of a specific variables. We will come back to this later.

```{r}
dbGetQuery(db, 
  "SELECT type, COUNT(type) AS type_count
  FROM posts
  GROUP BY type")
```

And we can then order the results using __ORDER BY__. The default is alphabetical order for the first column, but we can change that:

```{r}
# sort by type_count
dbGetQuery(db, 
  "SELECT type, COUNT(type) AS type_count
  FROM posts
  GROUP BY type
  ORDER BY type_count")

# now in descending orders
dbGetQuery(db, 
  "SELECT type, COUNT(type) AS type_count
  FROM posts
  GROUP BY type
  ORDER BY type_count DESC")

# which was the most popular post?
dbGetQuery(db, 
  "SELECT from_name, message, likes_count, datetime
  FROM posts
  ORDER BY likes_count DESC
  LIMIT 1")

```

## Querying multiple SQL tables

Now's where we get to see the magic of SQL. It makes it really easy to combine multiple SQL tables using __JOIN__. We just select columns from different tables and then specify what _key_ should be used to match them. The example belows shows an __inner join__, but there are other types (left vs right outer joins, cross joins...) that we will not cover today.

```{r}
dbGetQuery(db, 
  "SELECT posts.likes_count, congress.party, posts.date
  FROM posts JOIN congress
    ON congress.screen_name = posts.screen_name
  LIMIT 10")
```

## Grouping and aggregating

SQL makes it really easy to compute aggregate statistics of one column based on values of a different column, with a series of built-in functions (COUNT, AVG, MAX, MIN...). Let's now see a few examples of this type of queries:

```{r}
# use COUNT with * to return number of rows
dbGetQuery(db, 'SELECT COUNT(*) FROM posts')
dbGetQuery(db, 'SELECT COUNT(*) FROM congress')

# combining it with what we just learned
dbGetQuery(db, 
  "SELECT congress.party, COUNT(*)
  FROM posts JOIN congress
    ON congress.screen_name = posts.screen_name
  GROUP BY congress.party")

dbGetQuery(db, 
  "SELECT congress.party, congress.gender, COUNT(*)
  FROM posts JOIN congress
    ON congress.screen_name = posts.screen_name
  GROUP BY congress.party, congress.gender")
```


```{r}
# other functions: average
dbGetQuery(db, 
  "SELECT congress.party, AVG(posts.likes_count), COUNT(*) 
  FROM posts JOIN congress
    ON congress.screen_name = posts.screen_name
  GROUP BY congress.party")

# and max
dbGetQuery(db, 
  "SELECT congress.party, MAX(posts.likes_count), COUNT(*) 
  FROM posts JOIN congress
    ON congress.screen_name = posts.screen_name
  GROUP BY congress.party")

# and sum
dbGetQuery(db, 
  "SELECT congress.party, SUM(posts.likes_count)/COUNT(*) AS average 
  FROM posts JOIN congress
    ON congress.screen_name = posts.screen_name
  GROUP BY congress.party")
```






