---
title: "Challenge 2: Querying an Google BigQuery database"
author: Pablo Barbera
date: "August 11, 2017"
output: html_document
---

You will now continue practicing how to write SQL queries using a Google BigQuery database that contains all Citi Bike trips (NYC's bike-sharing service) since Citi Bike launched in September 2013. More information about the database is available [here](
https://cloud.google.com/bigquery/public-data/nyc-citi-bike).

First, connect to the database and count how many rows it contains:

```{r}
library(bigrquery)
project <- "usc-barbera"

get_table(project="bigquery-public-data",
          dataset="new_york",
          table="citibike_trips")

# how many trips in database?
query_exec(
  "SELECT COUNT(*) AS count
  FROM [bigquery-public-data:new_york.citibike_trips]",
	project = project, useLegacySql = FALSE)

# see first line
query_exec(
  "SELECT *
  FROM [bigquery-public-data:new_york.citibike_trips]
  LIMIT 1",
	project = project, useLegacySql = FALSE)
```

Which are the 10 most popular stations in terms of how many trips started there?

```{r}

```

What is the average trip duration based on the hour of the day when they start?

```{r}

```

What is the average trip duration of a CitiBike trip in NYC?

```{r}

```

What is the average trip duration based on the age of the riders?

```{r}

```

What is the average distance of a CitiBike trip in NYC?

```{r}

```

What is the average distance of a trip based on gender?

```{r}

```

What is the average speed of a CitiBike trip?

```{r}

```

What is the average speed based on age?

```{r}

```

