---
title: "Challenge 9: Supervised machine learning"
author: Pablo Barbera
date: "August 9, 2018"
output: html_document
---

Now it's your turn to build a classifier! The file `incivility.csv` contains roughly a stratified random sample of 3,000 Facebook comments on the public pages of U.S. legislators. Each of these comments was labeled by a human annotator as containing `attacks` or not (`atacks` here is defined as: name-calling, insults, aspersions..., etc). We will take this as our working definition of incivility.

Read the dataset into R. What proportion of comments in this sample are classified as being uncivil?

```{r}
library(quanteda)
fb <- read.csv("~/data/incivility.csv", stringsAsFactors = FALSE)

```

Create a DFM for the comments on this dataset. Use sensible options (e.g. should you stem? trim it?)

```{r}

```

Now, split the dataset into training and test sets using similar code as what we just used.

```{r}

```

Build a LASSO classifier with the training set.

```{r}

```

Compute the accuracy, precision, and recall. Do you think this is a good classifier?

```{r}


```

Find the features that are most predictive of each category. Do you think your classifier properly captured the concept of incivility?

```{r}


```

Try with one of the other classifiers we covered in class. Do you find that the performance of your classifier improves?

```{r}


```


(Advanced) Do uncivil comments receive higher engagement? Use the `comments_likes_count` variable (measuring the number of likes on each comment) to answer this question. Use any statistical technique you may consider appropriate.

```{r}


```




