---
title: "Challenge 5: Querying an Google BigQuery database"
author: Pablo Barbera
date: August 7, 2018
output: html_document
---

Now it's your turn to practice querying this BigQuery database! Write code to answer the following questions...

```{r}
library(bigrquery)
project <- "usc-barbera"
set_service_token("~/credentials/bigquery-token.json")
```

1. We will start with a follow-up to the example on validating the predicted party labels. Compute the proportion of users within each predicted party group that mentioned the words `progressive`, `liberal` or `democrat` in their description. Which group has the highest proportion?

```{r}


```

2. Compute the total number of tweets per hour of the day that mention the string `good morning`. Clue: you can use the `HOUR` operator. Then plot the count of tweets (y-axis) per hour of the day (x-axis). What do you find? Do you notice any interesting pattern?

```{r}


df <- df[order(df$hour),]
plot(df$hour, df$morning_count, type="l")
```

3. Now replicate this analysis but for tweets that mention the string `good night`. 

```{r}


```

4. To try to investigate why you find any surprising pattern, do a `JOIN` of the tweet-level table and the user-level table and look at a few of the `location` fields for users that tweet `good morning` or `good night` at unexpected times. What do you find?

```{r}


```

5. It was recently discovered that the Twitter acount `@TEN_GOP`, which appeared to belong to the Republican party of Tennessee, was actually handled by a Russian troll. This user is not in the dataset, but we can find out if any user in the sample retweeted any message by this account.

Create a database with all the retweets of this account. Clue: you can just search for the string `RT @TEN_GOP`. How many retweets are there?

```{r}



```

6. To get a sense of what this account was tweeting, use the code below to create a wordcloud of the most frequent words. Note that this code assumes you have a data frame called `df` with a variable named `text` that contains the text of the tweets.

```{r}
library(quanteda)
text <- paste(df$text, collapse=" ")
textplot_wordcloud(dfm(corpus(text),
                       remove_punct=TRUE,
                       remove=c(stopwords("en"), "@ten_gop", "rt", "https",
                                "t.co")),
                  max.words=100, scale=c(3, .5), rot.per=0)
```

7. Which party group was more likely to retweet this account? What about the age groups? 

```{r}





```

8. (Optional) Can you compute the proportion of users within each party that retweeted `@TEN_GOP`?


```{r}

```
